Quotients of the magmatic operad:\\ lattice structures and convergent rewrite systems

### 2018-09-11 Mag_quotients.tex
MAJ

We define the \Def{2-nihilpotente operad} by
\begin{equation}
  \label{equ:2-nihilpotente_operad}
  \NiHil:=\KMag/_{\INiHil}
  \end{equation}
where $\INiHil$ is the ideal spanned by the two trees $\LComb{3}$ and
$\RComb{3}$.
\Todo{Est-ce qu'il faut rajouter le 2 dans la macro désignant cette
  opérade ? De plus, Samuele,\\
  pourrais-tu rajouter la référence de Loday ?}
We easily check that we have
\begin{equation}
  \label{equ:Hilbert_series_of_NiHil}
  \HilbertSeries_{\NiHil}(t)=t+t^2.
\end{equation}
\Todo{Est-ce bien la bonne série ? Samuele, tu avais l'air de dire que
  c'était plutôt $t+2t^2$}

@@

\begin{proof}
  Let $R^{(\gamma)}$ be the set of rewrite rules
  $\Tfr\Rew\RComb{\gamma}$, where $\Tfr$ runs over all the binary trees
  of arity $\gamma+1$ different from $\RComb{\gamma}$. The unique normal
  form of arity $n\geq\gamma+1$ for  $R^{(\gamma)}$ is equal to
  $\RComb{n}$, so that $R^{(\gamma)}$ is a convergent presentation of
  \begin{math}
    \KMag/_{\IRC{\gamma}}.
  \end{math}
  Moreover, the normal forms of arity $n\leq\gamma$ are all the trees of
  arity $n$ and, by using the convergent presentation
  $R^{(\gamma)}$, we see that the compositions of
  \begin{math}
    \KMag/_{\IRC{\gamma}}
  \end{math}
  satisfy~\eqref{equ:compositions_of_RC}. Hence, $R^{(\gamma)}$ is also a
  convergent presentation of
  \begin{math}
    \RC{\gamma}
  \end{math}
  which proves Lemma~\ref{lem:presentation_of_RC}.
\end{proof}

@@

\smallbreak
Now, we show~\eqref{equ:example_upper-bound}. We denote by
\begin{math}
  \pi:\KMag\to\RC{3}
\end{math}
the natural projection. Let $\Tfr$ be a tree of arity $3$. From
Lemma~\ref{lem:presentation_of_RC}, the kernel of $\pi$ is spanned by
the elements $\alpha_{\Tfr}:=\Tfr-\RComb{3}$, where $\Tfr$ is a binary
tree of arity $4$. The elements $\alpha_{\Tfr}$ belong to the ideal
$\IAs$ of relations of $\KAs$ and to the ideal $\IAAs$ since both
$[\Tfr]_{\IAAs}$ and $[\RComb{3}]_{\IAAs}$ are equal to $[0]_{\IAAs}$
from Lemma~\ref{lem:properties_of_AAs}. Hence, the kernel of $\pi$ is
included in $\IAs\cap\IAAs$, so that, by definition of the upper-bound
$\SupQMag$, $\pi$ induces a surjective operad morphism
\begin{math}
  \overline{\pi}:\As\SupQMag\AAs\to\RC{3}
\end{math}.
We conclude using a dimension argument: by definition, the Hilbert series
of $\RC{3}$ is equal to
\begin{equation}
  \label{equ:Hilbert_series_of_RC3}
  t+t^2+2t^3+\sum_{n\geq 4}t^n
\end{equation}
and from the Grassmann formula analog given in
Theorem~\ref{thm:Grassmann_formula_for_Hilbert_series_of_QMag}, we check
that the Hilbert series of $\As\SupQMag\AAs$ is also equal
to~\eqref{equ:Hilbert_series_of_RC3}. Hence, $\overline{\pi}$ is an
isomorphism, which proves~\eqref{equ:example_upper-bound}.
  \end{proof}
\medbreak

>>>>

We define the \Def{2-nihilpotente operad} by
\begin{equation}
  \label{equ:2-nihilpotente_operad}
  \NiHil:=\KMag/_{\INiHil}
  \end{equation}
where $\INiHil$ is the ideal spanned by the two trees $\LComb{3}$ and
$\RComb{3}$.
\Todo{Est-ce qu'il faut rajouter le 2 dans la macro désignant cette
  opérade ? De plus, Samuele,\\
  pourrais-tu rajouter la référence de Loday ?}
We have
\begin{equation}
  \label{equ:Hilbert_series_of_NiHil}
  \HilbertSeries_{\NiHil}(t)=t+t^2.
\end{equation}
\Todo{Est-ce bien la bonne série ? Samuele, tu avais l'air de dire que
  c'était plutôt $t+2t^2$. Ou peut-être que NiHil est la $1$-
  nihilipotente opérade ?}

@@

\begin{proof}
  Let $R^{(\gamma)}$ be the set of rewrite rules
  $\Tfr\Rew\RComb{\gamma}$, where $\Tfr$ runs over all the binary trees
  of arity $\gamma+1$ different from $\RComb{\gamma}$. The unique normal
  form of arity $n\geq\gamma+1$ for  $R^{(\gamma)}$ is equal to
  $\RComb{n}$, so that $R^{(\gamma)}$ is a convergent presentation of
  \begin{math}
    \KMag/_{\IRC{\gamma}}.
  \end{math}
  Moreover, the normal forms of arity $n\leq\gamma$ are all the trees of
  arity $n$ and, by using the convergent presentation
  $R^{(\gamma)}$, the compositions of
  \begin{math}
    \KMag/_{\IRC{\gamma}}
  \end{math}
  satisfy~\eqref{equ:compositions_of_RC}. Hence, $R^{(\gamma)}$ is also a
  convergent presentation of
  \begin{math}
    \RC{\gamma}
  \end{math}
  which proves Lemma~\ref{lem:presentation_of_RC}.
\end{proof}

@@

  \smallbreak
  Now, we show~\eqref{equ:example_upper-bound}. We denote by
  \begin{math}
    \pi:\KMag\to\KAs\SupQMag\AAs
  \end{math}
  the natural projection. Let $\Tfr$ be a tree of arity $4$. From
  Lemma~\ref{lem:presentation_of_RC}, the kernel of $\pi$ is spanned by
  the elements $\alpha_{\Tfr}:=\Tfr-\RComb{3}$, where $\Tfr$ is a binary
  tree of arity $4$. The elements $\alpha_{\Tfr}$ belong to the ideal
  $\IAs$ of relations of $\KAs$ and to the ideal $\IAAs$ since both
  $[\Tfr]_{\IAAs}$ and $[\RComb{3}]_{\IAAs}$ are equal to $[0]_{\IAAs}$.
  The last statement is shown using the convergent
  presentation~\eqref{equ:convergent_presentation_of_AAS}. Hence, the kernel
  of $\pi$ is included in $\IAs\cap\IAAs$, so that, by definition of the
  upper-bound $\SupQMag$, $\pi$ induces a surjective operad morphism
  \begin{math}
    \overline{\pi}:\As\SupQMag\AAs\to\RC{3}
  \end{math}.
  We conclude using a dimension argument: by definition, the Hilbert series
  of $\RC{3}$ is equal to
  \begin{equation}
    \label{equ:Hilbert_series_of_RC3}
    t+t^2+2t^3+\sum_{n\geq 4}t^n
  \end{equation}
  and from the Grassmann formula analog given in
  Theorem~\ref{thm:Grassmann_formula_for_Hilbert_series_of_QMag}, we check
  that the Hilbert series of $\As\SupQMag\AAs$ is also equal
  to~\eqref{equ:Hilbert_series_of_RC3}. Hence, $\overline{\pi}$ is an
  isomorphism, which proves~\eqref{equ:example_upper-bound}.
\end{proof}
\medbreak
